Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-04 paper_jasa.tex
experiment 3 details

This experiment is designed for the same regression model (Equation \ref{multi}) with $p=1$ and covariate $X_1$ being continuous. $X_1$ is simulated from standard normal(0,1). Data is simulated with two sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The different values for slope parameter that are used to generate data are shown in Table \ref{tbl:experiment_params}. We arbitrarily fix $\beta_0 = 6$ while generating actual data. For each combination of parameter, we generate at least three different actual data sets constituting a total of 65 lineups. Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots.

@@

@@

\[
  Y_i = \left\{
  \begin{array}{l l}
    5+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad \text{ $i$ =1,...,85}\\
    15+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=86,...,n\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$. One such plot of the contaminated data for $\sigma$=5 and $\beta$=5 is shown in figure \ref{fig:cont_dat}. For this experiment we only consider sample size $n=100$ and error standard deviation $\sigma=5$.

For this experiment we generate five replications of actual data sets for each slope value shown in table \ref{tbl:experiment_params}. This produces 30 lineups.

\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects. Show a sample lineup, probably the most difficult, instead of the single plot.}

>>>>

This experiment is designed for the same regression model (Equation \ref{multi}) with $p=1$ and covariate $X_1$ being continuous. $X_1$ is simulated from standard normal(0,1). Data is simulated with two sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The different values for slope parameter that are used to generate data are shown in Table \ref{tbl:experiment_params}. We arbitrarily fix $\beta_0 = 6$ while generating actual data. For each combination of parameter, we generate at least three different actual data sets constituting a total of 65 lineups. For some parameter combinations we have more than three replications. Unlike what we did in section \ref{sec:category} we did not use any blocking structure to simulate actual data and null plots.

@@

{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation. For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in table \ref{tbl:experiment_params}. For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4. The 10 lineups are choosen as follows;

\begin{enumerate}
	\item For one combination of  sample size ($n$) and error SD ($\sigma$) randomly select five lineups with one from each difficulty level
	\item For another combination of sample size ($n$) and error SD ($\sigma$) select four lineups with one from each difficulty level 0 through 3
	\item Finally select another lineup with difficulty level 0
\end{enumerate}

@@

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=n+1, ...,n*c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data. For the experiment we consider $n=100$, $c=0.15$ which produced actual data size 115, $\alpha=0$ and $\lambda=10$ . $\sigma$ is chosen such that overall error standard deviation of the data is 5 which is approximately 3.5. One such plot of the contaminated data for  $\beta$=5 is shown in figure \ref{fig:cont_dat}.  We fit model (Equation \eqref{multi}) with $p=1$  and intercept $\beta_0=0$ to this contaminated data to test the hypothesis whether slope is zero or not.

{\bf Steps to obtain lineup}
\begin{enumerate}
	\item For a given parameter setting obtain actual data set
	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
	\item Generate actual data plot (a scatter plot of Residual vs $X_1$)
	\item Generate 19 null residual data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
\end{enumerate}

One lineup for slope $\beta=0.4$ is shown in figure \ref{fig:test_contaminated}. For this experiment we generate five replications of actual data sets for each slope value shown in table \ref{tbl:experiment_params}. This produces 30 lineups. each subject got at least 10 lineups to evaluate. Based on the slope parameter we group lineups into six groups with difficulty = 0 for largest slope and difficulty = 5 for the smallest slope. Eight (8) lineups with two from each of the difficulty group from 0 through 3 and two(2) lineups with one from each difficulty group 4 and 5 were randomly picked for each subject.

\begin{figure}[htp]
%\begin{figurehere}
   \centering
       \includegraphics[width=0.95\textwidth]{lineup_contaminated.pdf}
       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous but the inclusion of some contamination with the data spoils the normality assumption of error structure. One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true. Which plot is the most different from the others, in the sense that there is the steepest slope?}
       \label{fig:test_contaminated}
\end{figure}

%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects. Show a sample lineup, probably the most difficult, instead of the single plot.}
